On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-02-28 manuscript.Rmd
fixed comparison and description thereof. Nearly there!

\
In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter. In this context, it is useful to briefly summarize the approach taken by the majority of vegetation model. Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq. 1).

@@

\
Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge. Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq. 1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate the large acclimatory potential of this trait to environmental drivers.

@@

| Variable      |  Definition                                                                                       |
| ------------- | ------------------------------------------------------------------------------------------------- |
| g~cuti~       | Conductance of an isolated non-stomatal cuticle.                                                  |
| g~min~        | Minimum conductance measured from weight loss of detached leaves.                                 |
| g~dark~       | Night-time conductance, or conductance after significant dark adaptation.                         |
| g~0~          | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
| g~s,low A~    | Cuvette measurements of g~s~ when photosynthesis rates are very low.                              |

@@

\
We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data. To estimate g~0~ (Eq. 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation). All conductance values were converted to per unit projected surface area. Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the 52 studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted or extrapolated g~0~ term.

@@

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark (further discussed in Section VI). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.

\
The fitted minimum conductance parameter, g~0~ (mean = 32.8 mmol m^-2^ s^-1^), was much higher than g~min~, but similar to g~dark~. However, we point out the uncertainties associated with g~0~ in Section III, which tend to inflate estimates of g~0~. It is telling that actual measured stomatal conductance during low photosynthesis (g~s,low A~) was  much lower than g~0~ or g~dark~. In other words, actual daytime conductance during these conditions that cause low photosynthetic rates is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their Fig. A2).

@@

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq. 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

@@

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with the photon flux density (PPFD), as is typical in leaf gas exchange data. When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

Eq. 3 - derived using zero g~0~ - thus does not give any dependence on PPFD.

@@

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition when net photosynthesis is zero. However as our literature comparison showed (Fig. 1), g~dark~ is substantially higher compared to all other variables except fitted g~0~. The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

\
First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~, and should be modelled in a separate framework that is yet to be identified.

@@

\
**Figure 8**. Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 °C) [see @drake2018]. Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, mixed-effects model).

@@

\
In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression, the very high values compared to minimum conductance reached for detached leaves, we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models. Then, how should g~0~ be estimated? It is clear from our review and synthesis of available data that there is no single minimum conductance. Leaves maintain much higher g~dark~ (itself an actively controlled process) than the minimum conductance measured on intact detached leaves. Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants.

\
Our first conclusion is that g~0~, g~min~ and g~dark~ should not be assumed equal in any modelling study. Indeed, a productive way forward is to view each of these conductances as distinct, and they should be parameterized separately. For example, when considering modelling stomatal conductance with Eq. 1, we suggest that g~min~ and g~0~ can be combined:

$$ g_s = min(g_{min}, g_0) + g_1 \frac{A_n}{C_a}f(D)$$

where g~min~ is parameterized with data obtained from mass loss of detached leaves (Fig. 6), representing a true whole-leaf minimum that is reached under conditions of severe drought, and g~0~ is the minimum reached during low photosynthesis (due to low light or other factors) for well-watered plants. During drought, g~0~ will decrease [@misson2004], so that the true minimum reached for Eq. 1 is g~min~.

\
It is however an open question on how g~min~ should be parameterized for different species, as we found both large variation between species that could not be explained by traits, and remarkable plasticity to growing conditions. Perhaps this plasticity is the reason that g~min~ did not vary predictably between species. Future studies should compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.

>>>>

\
In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter. In this context, it is useful to briefly summarize the approach taken by the majority of vegetation models. Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq. 1).

@@

\
Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge. Our goals are to a) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq. 1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.

@@

| Variable      |  Definition                                                                                       |
| ------------- | ------------------------------------------------------------------------------------------------- |
| g~cuti~       | Conductance of an isolated non-stomatal cuticle.                                                  |
| g~min~        | Minimum conductance measured from weight loss of detached leaves.                                 |
| g~dark~       | Night-time conductance, or conductance after significant dark adaptation.                         |
| g~0~          | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |

@@

\
We quantitatively compared each of the different 'minimum' conductance (Table 1) definitions by synthesizing existing data. Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation). All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented as such. We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought). Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca. 150 species during the night (we selected only gas exchange based measurements from their database). Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.

@@

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark (further discussed in Section VI). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks. In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime wither in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~). These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).

@@

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

@@

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data. When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

Eq. 3 - derived using zero g~0~ - thus does not give any dependence on PAR.

@@

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition when net photosynthesis is zero. We provide some caution to the assumption that g~dark~ can be used in models of daytime leaf conductance. First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~, and should be modelled in a separate framework that is yet to be identified.

@@

\
**Figure 8**. Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 °C) [see @drake2018]. Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model).

@@

\
In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models. Then, how should g~0~ be estimated? It is clear from our review and synthesis of available data that there is no single minimum conductance. Leaves maintain much higher g~dark~ (itself an actively controlled process) than the minimum conductance measured on intact detached leaves. Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants.

\
We suggest that the minimum conductance in the Ball-Berry model should include both a g~0~ and a g~min~ term, since g~0~ represents the minimum reached during low light and conditions of low photosynthesis, and g~min~ represents the minimum reached during severe drought. Moreover, we suggest that the minimum conductance is not simply added to the photosynthesis-dependent term (right-hand side in Eq. 1), but used as an actual minimum. Thus,

$$ g_s = max( min(g_{min}, g_0), g_1 \frac{A_n}{C_a}f(D))$$

\
This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see @misson2004]. Another advantage is that independent estimates of g~0~ can be used, and not with regression which produces the undesirable correlation with estimates of g~1~. The above formulation is yet to be tested against data, but we propose this test to be done combining data from drought and non-drought conditions.

\
Finally, we conclude that g~min~ displays a large amount of variation between species that could not be explained by traits, and remarkable plasticity to growing conditions. Perhaps this plasticity is the reason that g~min~ did not vary predictably between species. Future studies should compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.
